---
output: github_document
---

# DarwinShinyModules

[![Lifecycle:experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Build Status](https://github.com/darwin-eu/DarwinShinyModules/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu/DarwinShinyModules/actions?query=workflow%3AR-CMD-check)
[![codecov.io](https://codecov.io/github/darwin-eu/DarwinShinyModules/coverage.svg?branch=main)](https://codecov.io/github/darwin-eu/DarwinShinyModules?branch=main)

`DarwinShinyModules` is an R package containing shiny modules. Each module is
represented as an `R6` object, which contains pre-specified UI elements, and
back-end code.

## Features
- Provides pre-specified modules of common shiny app components.
- Provides pre-specified modules for DARWIN EU© analytical pipelines (in progress).
- Ability to overwrite UI elements and the server back-end with bespoke code.
- Compatible with bespoke shiny application development.
- Seamlessly integrate existing `OhdsiShinyModules`.
- Extend the existing library with your own modules

## Package support

| Package | Support |
| --- | --- |
| CohortCharacteristics | [![latest](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-latest.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-latest.yaml) [![1.1.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-1_1_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-1_1_0.yaml) [![1.0.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-1_0_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortCharacteristics-1_0_0.yaml) |
| CohortSurvival | [![latest](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-latest.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-latest.yaml) [![1.1.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-1_1_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-1_1_0.yaml) [![1.0.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-1_0_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/CohortSurvival-1_0_0.yaml) |
| IncidencePrevalence |  [![latest](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-latest.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-latest.yaml) [![1.2.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_2_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_2_0.yaml) [![1.1.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_1_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_1_0.yaml) [![1.0.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_0_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/IncidencePrevalence-1_0_0.yaml) |
| TreatmentPatterns | [![latest](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-latest.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-latest.yaml) [![3.1.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-3_1_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-3_1_0.yaml) [![3.0.0](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-3_0_0.yaml/badge.svg)](https://github.com/darwin-eu-dev/DarwinShinyModules/actions/workflows/TreatmentPatterns-3_0_0.yaml) |

## Installation
1. To install the latest stable version:

```{r install, eval=FALSE}
install.packages(remotes)
remotes::install_github("darwin-eu/DarwinShinyModules")
```


## Usage
```{r app, message=FALSE, eval=FALSE}
library(DarwinShinyModules)

irisTable <- Table$new(iris)
carsTable <- Table$new(mtcars)

launchDarwinDashboardApp(
  list(
    Tables = list(
      Iris = irisTable,
      MT_Cars = carsTable
    )
  )
)
```

## License
DarwinShinyModules is licensed under Apache License 2.0

