---
title: "Compatibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compatibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Compatibility
`DarwinShinyModules` is compatible with other modules or shiny code. One module that you might have noticed in the documentation is the `OhdsiModule`. The `OhdsiModule` is a module that wraps around the `xViewer` and `xServer` functions. This makes modules from `OhdsiShinyModules` compatible by default.

```{r eval=FALSE}
library(DarwinShinyModules)
library(DatabaseConnector)
library(ResultModelManager)
library(OhdsiShinyModules)
library(ShinyAppBuilder)

connectionDetails <- createConnectionDetails(
  dbms = "sqlite",
  server = "./results.sqlite"
)

connectionHandler <- ConnectionHandler$new(
  connectionDetails = connectionDetails
)

estimation <- OhdsiModule$new(
  connectionHandler = connectionHandler,
  viewerFun = estimationViewer,
  serverFun = estimationServer,
  resultDatabaseSettings = createDefaultResultDatabaseSettings()
)

preview(estimation)
```

If you want to create your own wrapper arround other existing modules, or shiny app code you can follow the examples in the *Creating A New Module* vignette.
