---
title: "Loading Data From Different Sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading Data From Different Sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
Data can come from anywhere. `DarwinShinyModules` assumes *you* load in the data from whatever source you like. This may be a csv-file, a table from a database. But in the end the data is loaded into memory in R. The intention behind this is that individual modules are (and should be) small. But can be used in large overarching modules.

In this vignette we will explore how to load in data from files and databases, by either pre-loading the data into memory. Or by loading data dynamically, during run time. Finally we will also show how to use the `DatabaseDBI` and `DatabaseDBC` modules.

## Preloading data
### From a File
```{r, eval=FALSE}
library(DarwinShinyModules)

tempDir <- file.path(tempdir(), "loading_data_example")
dir.create(tempDir, showWarnings = FALSE, recursive = TRUE)
write.csv(iris, file.path(tempDir, "iris.csv"), row.names = FALSE)
write.csv(mtcars, file.path(tempDir, "mtcars.csv"), row.names = FALSE)

data <- read.csv(file.path(tempDir, "iris.csv"))

tableFile <- Table$new(data = data)

if (interactive()) {
  preview(tableFile)
}
```

### A database using `DBI`
```{r, eval=FALSE}
library(DBI)
library(duckdb)
library(dplyr)

driver <- duckdb::duckdb(
  dbdir = file.path(tempDir, "database.duckdb")
)

con <- DBI::dbConnect(
  drv = driver
)

DBI::dbWriteTable(
  conn = con,
  name = "iris",
  value = iris,
  overwrite = TRUE
)

DBI::dbWriteTable(
  conn = con,
  name = "mtcars",
  value = mtcars,
  overwrite = TRUE
)

iris_db <- dplyr::tbl(src = con, "iris")

data <- iris_db %>%
  dplyr::collect()

tableDB <- Table$new(data = data)

if (interactive()) {
  preview(tableDB)
}
```

### From a database using `DatabaseConnector`
```{r, eval=FALSE}
library(DatabaseConnector)

connectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms = "sqlite", 
  server = file.path(tempDir, "database.sqlite")
)

connection <- DatabaseConnector::connect(connectionDetails)

DatabaseConnector::insertTable(
  connection = connection,
  databaseSchema = "main",
  tableName = "iris",
  data = iris
)

DatabaseConnector::insertTable(
  connection = connection,
  databaseSchema = "main",
  tableName = "mtcars",
  data = mtcars
)

data <- DatabaseConnector::renderTranslateQuerySql(
  connection = connection,
  sql = "SELECT * FROM iris;"
)

DatabaseConnector::disconnect(connection)

tableDBC <- Table$new(data)

if (interactive()) {
  preview(tableDBC)
}
```

## Using a database in your module
You can dynamically load data from files or a database in your own module. The loaded data can then be used in regular `DarwinShinyModule` modules.

The following examples show how to dynamically load from:

1. CSV-files
2. A database using `DBI` while maintaining the database connection
3. A database using `DatabaseConnector`, while managing the connection in run time.

### From csv-files
```{r, eval=FALSE}
TablesFromFile <- R6::R6Class(
  classname = "TablesFromFile",
  inherit = ShinyModule,

  public = list(
    initialize = function(filePath) {
      private$.filePath <- filePath
      # initialize the `Table` module with an empty data.frame
      private$.table <- Table$new(data = data.frame())
      private$.table$parentNamespace <- self$namespace
    }
  ),
  private = list(
    ## Fields
    .filePath = "",
    .table = NULL,

    ## Methods
    .UI = function() {
      shiny::tagList(
        # Select input to get csv-files from `tempDir`
        shiny::selectInput(
          inputId = shiny::NS(self$namespace, "file"),
          label = "File",
          # Get csv-file names for `tempDir`
          choices = private$getCsvFiles()
        ),
        private$.table$UI()
      )
    },

    .server = function(input, output, session) {
      private$.table$server(input, output, session)

      # Trigger on the file-selector
      shiny::observeEvent(input$file, {
        # Update the data field in the `Table` module with the selected csv-file
        private$.table$data <- read.csv(file.path(private$.filePath, input$file))
      })
    },

    # Get all csv-files from `tempDir`
    getCsvFiles = function() {
      allFiles <- basename(list.files(private$.filePath))
      allFiles[endsWith(allFiles, suffix = ".csv")]
    }
  )
)

mod <- TablesFromFile$new(tempDir)

if (interactive()) {
  preview(mod)
}
```

### From a database using `DBI`
In this example we maintain the connection to the database, post run-time.
```{r, eval=FALSE}
TablesFromDBI <- R6::R6Class(
  classname = "TablesFromDBI",
  inherit = ShinyModule,

  public = list(
    initialize = function(con) {
      private$.con <- con
      # initialize the `Table` module with an empty data.frame
      private$.table <- Table$new(data = data.frame())
      private$.table$parentNamespace <- self$namespace
    }
  ),
  private = list(
    ## Fields
    .con = NULL,
    .table = NULL,

    ## Methods
    .UI = function() {
      shiny::tagList(
        # Select input to get csv-files from `tempDir`
        shiny::selectInput(
          inputId = shiny::NS(self$namespace, "table"),
          label = "Table",
          # List tables from the connection
          choices = DBI::dbListTables(private$.con)
        ),
        private$.table$UI()
      )
    },

    .server = function(input, output, session) {
      private$.table$server(input, output, session)

      # Trigger on the file-selector
      shiny::observeEvent(input$table, {
        # Fetch table from DBI connection
        private$.table$data <- dplyr::tbl(src = private$.con, input$table) %>%
          # We need to collect, as the table needs to be in memory
          dplyr::collect()
      })
    }
  )
)

mod <- TablesFromDBI$new(con)

if (interactive()) {
  preview(mod)
}

DBI::dbDisconnect(con)
```

### From a database using `DatabaseConnector`
In this example we manage the connection to the database within run time.
```{r, eval=FALSE}
TablesFromDBC <- R6::R6Class(
  classname = "TablesFromDBC",
  inherit = ShinyModule,

  public = list(
    initialize = function(connectionDetails) {
      # Save connectionDetails
      private$.connectionDetails <- connectionDetails
      # initialize the `Table` module with an empty data.frame
      private$.table <- Table$new(data = data.frame())
      private$.table$parentNamespace <- self$namespace
    }
  ),

  private = list(
    ## Fields
    .connectionDetails = NULL,
    .connection = NULL,
    .table = NULL,

    ## Methods
    .UI = function() {
      shiny::tagList(
        # Select input to get csv-files from `tempDir`
        shiny::selectInput(
          inputId = shiny::NS(self$namespace, "table"),
          label = "Table",
          # Initialize empty, we update once the connection is established.
          choices = ""
        ),
        private$.table$UI()
      )
    },

    .server = function(input, output, session) {
      # Connect to the database if `.connection` is `NULL`
      if (is.null(private$.connection)) {
        private$.connection <- DatabaseConnector::connect(connectionDetails)
      }

      # Run the `finalize` method when the session ends >>this includes refreshing the session<<
      shiny::onStop(private$finalize)

      private$.table$server(input, output, session)
      
      # If we're connected to the database
      if (!is.null(private$.connection)) {
        # The `updated` flag ensures we update the choices once. Otherwise we
        # continuously query the database for table names. Which is not desirable
        # for this use case. The database has a separate state, and can therefore
        # change at will, in a different connection. New tables could appear,
        # existing tables could disappear.
        updated <- FALSE
        if (!updated) {
          # Update choices in select
          shiny::updateSelectInput(
            inputId = "table",
            choices = private$fetchTableNames()
          )
          updated <- TRUE
        }
  
        # Trigger on the file-selector
        shiny::observeEvent(input$table, {
          # Only query the table when the selector is updated
          if (input$table != "") {
            # Fetch table from the database
            private$.table$data <- DatabaseConnector::renderTranslateQuerySql(
              connection = private$.connection,
              sql = "SELECT * FROM @table",
              table = input$table
            )
          }
        })
      }
    },

    # Fetches the names of the tables in SQLite
    fetchTableNames = function() {
      if (!is.null(private$.connection)) {
        DatabaseConnector::renderTranslateQuerySql(
        connection = private$.connection,
        sql = "SELECT name FROM sqlite_master WHERE type = 'table'"
      ) %>%
          dplyr::pull(.data$NAME)
      }
    },

    # A feature of R6 is that the `Finalize` method also runs when the object is garbage collected.
    finalize = function() {
      if (!is.null(private$.connection)) {
        # Disconnect from database
        DatabaseConnector::disconnect(private$.connection)
        # Set `.connection` to `NULL`
        private$.connection <- NULL
      }
    }
  )
)

mod <- TablesFromDBC$new(connectionDetails)

if (interactive()) {
  preview(mod)
}
```

## Database Modules
These database modules were developed based on the previous example. Two implementations exist as of writing, a database module using `DatabaseConnector`, and a module using `DBI`.

The upside of using these modules is that they manage the connection for you. They open the connection on start-up, and close it on shutdown.

When the connection unexpectedly drops, it will also prompt the user to re-connect to the database.

### DatabaseDBI
The `DatabaseDBI` module uses `DBI` to interface with the database. This is ideal when you want to use `dplyr` syntax to query a database during run-time of your shiny app.

```{r, eval=FALSE}
# Setup the module with the driver for the database
db <- DatabaseDBI$new(driver)

# We initialize the table modules with an empty data.frame
irisTable <- Table$new(data.frame(), title = NULL)
carTable <- Table$new(data.frame(), title = NULL)

ui <- shiny::fluidPage(
  irisTable$UI(),
  carTable$UI(),
  db$UI()
)

server <- function(input, output, session) {
  # Run the server methods
  db$server(input, output, session)
  irisTable$server(input, output, session)
  carTable$server(input, output, session)

  # Attach tables `iris` and `mtcars` from the database
  db$attachTables("iris", "mtcars")

  # Update `data` field in `irisTable` module
  irisTable$data <- db$tables$iris %>%
    # We can use dplyr syntax that is supported by the driver
    dplyr::mutate(
      foo = .data$Sepal.Length * .data$Sepal.Width
    ) %>%
    head() %>%
    # We MUST collect the data, as `DarwinShinyModules` expects the final result
    # to be in memory.
    dplyr::collect()

  carTable$data <- db$tables$mtcars %>%
    dplyr::group_by(.data$cyl) %>%
    dplyr::summarise(mean_hp = mean(.data$hp, na.rm = TRUE), .groups = "drop") %>%
    dplyr::collect()
}

if (interactive()) {
  shiny::shinyApp(ui, server)
}
```

### DatabaseDBC
The `DatabaseDBC` module uses `DatabaseConnector` to interface with the database. This is ideal when you want to use SQL to query a database during run-time of your shiny app.

```{r, eval=FALSE}
# Setup the module with the ConnectionDetails of the database
db <- DatabaseDBC$new(connectionDetails)

# We initialize the table modules with an empty data.frame
irisTable <- Table$new(data.frame(), title = NULL)
carTable <- Table$new(data.frame(), title = NULL)

ui <- shiny::fluidPage(
  irisTable$UI(),
  carTable$UI(),
  db$UI()
)

server <- function(input, output, session) {
  # If we want to run code in shiny::onStop() we must do that before the
  # server() method.
  stopFun <- function() {
    db$execute(
      # Drop a table we create later
      sql = "DROP TABLE IF EXISTS tmp_table;"
    )
  }

  shiny::onStop(stopFun)

  # Run the server methods
  db$server(input, output, session)
  irisTable$server(input, output, session)
  carTable$server(input, output, session)

  # query the database, we can use all arguments from DatabaseConnectors
  #`renderTranslateQuerySql()`
  irisTable$data <- db$query(
    sql = "SELECT * FROM @schema.iris LIMIT 10",
    schema = "main"
  )

  # We can also execute queries in the database
  db$execute("
    DROP TABLE IF EXISTS tmp_table;
    
    SELECT
      cyl,
      AVG(hp) as avg_hp
    FROM @schema.mtcars
    GROUP BY cyl
    INTO tmp_table;",
    schema = "main"
  )

  # Assign the result from `tmp_table` to `carTable`
  carTable$data <- db$query(
    sql = "SELECT * FROM @schema.tmp_table;",
    schema = "main"
  )
}

shiny::shinyApp(ui, server)
```

