---
title: "(Re)-deploying Shiny Apps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{(Re)-deploying Shiny Apps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
DarwinShinyModules provides functions to (re-)deploy shiny apps from a `appStructure` object. An `appStructure` object is **named** a `list()` containing `ShinyModule` objects or nested lists containing `ShinyModule` objects. The structure of the `appStructure` list resembles how to navigate your shiny app, i.e:
```{r}
library(DarwinShinyModules)

# Make some modules to make the `appStructure` with
base <- Text$new("**base**")
nested_a <- Text$new("**nested A**")
nested_b <- Text$new("**nested B**")
sub_a <- Text$new("**sub A**")
sub_b <- Text$new("**sub B**")
comb_a <- Text$new("**comb A**")
comb_b <- Text$new("**comb B**")
comb_c <- Text$new("**comb C**")

# Creating the `appstructure`
appStructure <- list(
 base = base,
 nested = list(nested_a, nested_b),
 nested_sub = list(
   sub_a = sub_a,
   sub_b = sub_b
 ),
 nested_combined = list(
   comb_a_b = list(comb_a, comb_b),
   comb_c = comb_c
 )
)
```

The main `list()` is always named, as for your navigation panel or bar, each items should have a name. Nested lists (see `appStructure$nested`) do not have to be named. The modules will be put underneath each other on the same page. If you do name the nested list (see `appStructure$nested_sub`), sub-items will be created for the main menu item.

## Deploying from an `appStructure`
We can deploy a shiny app directly from an `appStructure` object. We can do this by calling `deployAppStructure()`
```{r, eval=FALSE}
deployAppStructure(
  appStructure = appStructure,
  appDir = tempdir())
```

`deployAppStructure()` is basically a wrapper around `rsconnect::deployApp()`. `deployAppStructure()` has the dot-dot-dot parameter (`...`) that will pass any other parameters you specify to `rsconnect::deployApp()`. This allows you to pass things like an `appId` or account arguments.

## Re-deployment
Lets say we would like to restructure our shiny app based on some feedback we have gotten. We can download a tar.gz file from posit. This file is structured as followed:

```
bundle-xxxx.tar.gz
  |- .
    |- <...>
    |- appStructure.qs
```

There is no need to unpack the tar.gz-file, `loadFromDisk()` will extract the `appStructure` object for us:
```{r}
appStructure <- loadAppStructure(
  filePath = "./bundle-xxxx.tar.gz",
  appStructureFileName = "appStructure.qs"
)
```

Now we have an identical copy of the `appStructure` that we made before. We can alter the `appStructure` as we like:
```{r}
# Remove `base`
appStructure$base <- NULL

# Re-order alphabetically:
appStructure <- appStructure[sort(names(appStructure))]
```

With our updates made, we can re-deploy our app, by using the `deployAppStructure()` function again, note that we pass the `appId` argument, which is found in the `info` tab on posit:
```{r, eval=FALSE}
deployAppStructure(
  appStructure = appStructure,
  appDir = tempdir(),
  appId = 123)
```
