---
title: "An AppStructure object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An AppStructure object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
In examples, documentation, and code you might have seen a reference to an `appStructure`. An `appStructure` object is **named** a `list()` containing `ShinyModule` objects or nested lists containing `ShinyModule` objects. The structure of the `appStructure` list resembles how to navigate your shiny app, i.e:
```{r}
library(DarwinShinyModules)

# Make some modules to make the `appStructure` with
base <- Text$new("**base**")
nested_a <- Text$new("**nested A**")
nested_b <- Text$new("**nested B**")
sub_a <- Text$new("**sub A**")
sub_b <- Text$new("**sub B**")
comb_a <- Text$new("**comb A**")
comb_b <- Text$new("**comb B**")
comb_c <- Text$new("**comb C**")

# Creating the `appstructure`
appStructure <- list(
 base = base,
 nested = list(nested_a, nested_b),
 nested_sub = list(
   sub_a = sub_a,
   sub_b = sub_b
 ),
 nested_combined = list(
   comb_a_b = list(comb_a, comb_b),
   comb_c = comb_c
 )
)
```

The main `list()` is always named, as for your navigation panel or bar, each items should have a name. Nested lists (see `appStructure$nested`) do not have to be named. The modules will be put underneath each other on the same page. If you do name the nested list (see `appStructure$nested_sub`), sub-items will be created for the main menu item.
